### 2017-05-29 distr-03-design.tex
Wording, fit with previous intro into distr

The SG-MCMC a-MMSB algorithm has an abundance of
opportunities for parallelism, for the multi-threaded shared-memory type as
well as for distributed-memory parallelism. The benefit of multi-threaded
parallelism is the speedup of the computation. Additionally, a distributed implementation
allows the storage of data in the collective memory of a cluster
and increases aggregate memory bandwidth. Both of these parameters scale with the number of
machines. The downside of a distributed implementation is that it requires
considerable implementation effort.

@@

This section describes how we created a distributed version of the C++
algorithm. We didn't use the advanced intra-node parallelization techniques
described in the previous section, but found that straightforward
multi-threaded parallelization techniques sufficed: the distributed performance
is dominated by (remote) memory latency, not by node parallelism. As our
analysis will show, our pipelining optimizations will fully overlap the
parallel calculations with (pre)loading the data from remote hosts.
The details of the multi-threaded parallelism
will be discussed only where appropriate. The
distributed design follows a master-worker paradigm, where
the master controls the sequence of parallel operations and the workers perform the
calculations. For thread parallelism, we annotate the program with
OpenMP~\cite{OpenMPSpec}. For distributed communication, we use
MPI~\cite{Forum:1994:MMI:898758}.

>>>>

The implementations in the previous section put a limit to the size of
community networks and number of communities that can be processed: the
state must fit into GPU memory. To tackle networks of much larger size, we
created a distributed implementation that stores the data in the
collective memory of a cluster. The downside of a distributed implementation
is that the burden of memory access is increased for our already data-dominant
algorithm.

@@

This section describes how we created a distributed version of the C++
algorithm. We used the multicore CPUs in the DAS5 with straightforward
intra-node thread parallelisation, not the GPU implementations described
in the previous section. The justification is twofold: in our DAS5 system,
it is pointless to speed up the intra-node performance beyond straightforward
multithreading, because distributed performance is dominated by remote memory
latency; the analysis in Section~\ref{eval-pipeline} shows that the time for
(pre)loading the data from remote hosts greatly exceeds the calculation time.
The calculation time is fully overlapped in our pipelining implementation.
Moreover, usage of GPUs in the distributed setting would necessitate staging of
the relevant parts of $\pi/\phi$ from the host to GPU memory in each iteration,
which would eat most of the performance gain offered by the GPU because
the algorithm is so data-intensive.  % The details of the multi-threaded
parallelism % will be discussed only where appropriate.  The distributed design
follows a master-worker paradigm, where the master controls the sequence of
parallel operations and the workers perform the calculations. For thread
parallelism, we annotate the program with OpenMP~\cite{OpenMPSpec}. For
distributed communication, we use MPI~\cite{Forum:1994:MMI:898758}.
